---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# timerip

<!-- badges: start -->
[![R build status](https://github.com/DavisVaughan/timerip/workflows/R-CMD-check/badge.svg)](https://github.com/DavisVaughan/timerip)
[![Travis build status](https://travis-ci.org/DavisVaughan/timerip.svg?branch=master)](https://travis-ci.org/DavisVaughan/timerip)
[![Codecov test coverage](https://codecov.io/gh/DavisVaughan/timerip/branch/master/graph/badge.svg)](https://codecov.io/gh/DavisVaughan/timerip?branch=master)
<!-- badges: end -->

> In physical cosmology, the Big Rip is a hypothetical cosmological model concerning the ultimate fate of the universe, in which the matter of the universe, from stars and galaxies to atoms and subatomic particles, and even spacetime itself, is progressively torn apart by the expansion of the universe at a certain time in the future.

The goal of timerip is to be a fast extractor of individual date-time components, such as year or month, from `Date` or `POSIXct` objects.

## Installation

You can install the dev version from GitHub with:

``` r
devtools::install_github("DavisVaughan/timerip")
```

## Example

The main goal is to be faster than `as.POSIXlt()`, which wastes memory and time when you just want a single component.

```{r, eval = FALSE}
x <- as.POSIXct("2019-01-01", tz = "EST")
x <- rep(x, 2e5)

bench::mark(
  rip_year(x),
  as.POSIXlt(x)$year + 1900L
)
```

```{r, eval = FALSE}
#> # A tibble: 2 x 6
#>   expression                      min   median `itr/sec` mem_alloc `gc/sec`
#>   <bch:expr>                 <bch:tm> <bch:tm>     <dbl> <bch:byt>    <dbl>
#> 1 rip_year(x)                  5.29ms   6.39ms     153.      2.3MB     2.07
#> 2 as.POSIXlt(x)$year + 1900L   12.5ms  15.55ms      63.7    9.92MB    20.3
```

Extraction for `Date`s is faster too.

```{r, eval = FALSE}
y <- as.Date(x)

bench::mark(
  rip_year(y),
  as.POSIXlt(y)$year + 1900L
)
```

```{r, eval = FALSE}
#> # A tibble: 2 x 6
#>   expression                      min   median `itr/sec` mem_alloc `gc/sec`
#>   <bch:expr>                 <bch:tm> <bch:tm>     <dbl> <bch:byt>    <dbl>
#> 1 rip_year(y)                  4.92ms   5.83ms     162.    790.7KB     2.08
#> 2 as.POSIXlt(y)$year + 1900L  17.82ms  19.45ms      50.9    9.16MB    12.7
```
